Solution of an Acoustic Transmission Inverse Problem by Extended Inversion

### 2020-09-27 repro/papers/discrep/paper.tex
interim

This paper examines one of a number of proposed alternatives to least squares
FWI, based on an extension of the model-to-data mapping.  We show
exactly how an example of this {\em extended inversion} approach leads
to successful solution of a very simple acoustic transmission inverse
problem. Despite its simplicity, this problem exhibits several
important characteristics of all inverse problems in wave propagation
in which wave velocities are amongst the unknowns. It has the
advantage of permitting explicit calculations that reveal clearly the
failure mechanism for FWI, that is, straightforward nonlinear least
squares, and also the reasons for success of an extended inversion
approach.

In the remainder of this introduction, we will briefly describe the conceptual
framework, the inverse problem, and sketch our main results. We leave out
some technical details that are supplied in the next section,
containing precise statements of the results.

The idealized physical setting is the propagation of wave through a
homogeneous acoustic fluid occuyping $\bR^3$ \cite[]{Frie:58}. The
fluid's salient characteristic for the present discussion is the
slowness, or reciprocal wave velocity, $m > 0$. A source of acoustic
energy localized at a point in $\bR^3$ radiates uniformly in all
directions with time-dependent intensity $w(t)$. Both the acoustic
pressure field and the intensity (``wavelet'') $w$ vanish at large
negative time. A receiver records the pressure values at a distance
$r>0$ from the source position, between the times $t_{\rm min}$ and
$t_{\rm max}$. As explained in the next section, the
recorded presssure is predicted to be
\begin{equation}
\label{eqn:mod}
\frac{1}{4\pi r}w\left(t-mr\right)  = F[m]w(t)
\end{equation}
Ignoring the amplitude factor $1/(4\pi r)$, this map implements a $m$-dependent time
shift. This time shift operator is the basis of many descriptions of
the cycle-skipping phenomenon (for example, \cite{VirieuxOperto:09},
Figure 7), so it is unsurprising that an analysis of cycle-skipping
can be based on the simple modeling operator described above.

Evidently if the wavelet $w$ is square-integrable, so is the predicted
recorded pressure - its $L^2$ norm is the cumulative power trasferred
from the fluid to the receiver \cite[]{SantosaSymes:00}. Thus this relation defines a map
$F: \bR^+ \times L^2(\bR) \rightarrow
L^2([t_{\rm min},t_{\rm max}])$, linear in its second argument.

A naive version of the inverse problem investigated in this paper
would be: given recorded data
$d \in L^2([t_{\rm min},t_{\rm max}])$, find $m \in \bR$ and $w \in L^2(\bR)$ so that
$F[m]w \approx d$. Clearly this problem statement is not interesting as it
stands, since it is always possible to solve it with no error and nonuniquely,
and it does not constrain the slowness $m$ at all. To be useful, the statement must
be augmented with a constraint on the model $(m,w)$. We choose 
$\lambda > 0$ and search for $w$ with $\mbox{supp }w \subset
[-\lambda,\lambda]$. The quantity $\lambda$ is to be considered
``small'': as we shall show, it is related to a mean wavelength of
$w$.

@@

Despite its simplicity, this problem exhibits the
cycle-skipping pathology characteristic of field-scale FWI. Specifically, we show that the reduced mean square
error function, the objective function of FWI for this problem, 
\begin{equation}
  \label{eqn:redms}
  (m,w) \mapsto \lerr[m,w,d]=\frac{1}{2}\|\lF[m]w-d\|^2,
\end{equation}
has local minimizers far from any global minimizer, even for noise-free
data (Theorem \ref{thm:fwi}).

This fact implies that local optimization must fail to solve a precise
version of the FWI problem: given $\epsilon, \lambda > 0$, $(m,w)$ is
a solution if it lies in the domain of $\lF$ and
$\lerr[m,w,d] \le 1/2 \epsilon^2\|d\|^2$ (Problem Statement
\ref{eqn:probstat0} below).  In fact, we show
that``large residual'' local minima of $\lerr[\cdot,\cdot,d]$ exist,
at which its value is much greater than its global minimum
value. Therefore if $\epsilon$ is near the global minimum value, then
local minimization of $\lerr[\cdot,\cdot,d]$ can fail to find a
solution of the inverse problem. This phenomenon is what is meant by
``cycle-skipping''.

Extended inversion involves an extension of the model-to-data map, in
this case $\lF$.  The extension used here consists in dropping the
support constraint on $w$, so the extended map is simply $F$ as
defined above ($\lF$ is a restriction of $F$, so $F$ is an extension
of $\lF$). To compensate for the resulting indeterminacy in the
estimation of $m$, we add a penalty term to the objective: for $m \in
\bR^+,w\in L^2(\bR), d \in L^2([t_{\rm min},t_{\rm max}])$, define
\begin{eqnarray}
  \label{eqn:pen}
  e[m,w;d] & = & \frac{1}{2}\|F[m]w-d\|^2,\\
  g[w] & = & \frac{1}{2}\|A\|^2, \\
  \Ja[m, w;d] &=& e[m,w;d] + \alpha^2 g[w] \nonumber\\
  &=& \frac{1}{2}(\|F[m]w-d\|^2 + \alpha^2 \|Aw\|^2). 
\label{eqn:pen}
\end{eqnarray}
The only obvious requirement on the penalty operator $A$ is that it be
nonzero for large $|t|$, so that minimizers of $\Ja$ should be
concentrated near $t=0$. Thus the penalty is a soft stand-in for the
support constraint. We
show that a good choice for $A$ is given by $(Aw)(t)=tw(t)$, and the
remainder of the discussion in this section makes this assumption. The
penalty weight $\alpha$ must be chosen somehow; we review a simple
algorithm for controlling $\alpha$ later in the paper.

Assume that $\mu>0$ and $d = F[m_*]w_* + n$, with $\mbox{supp }w_*
\subset [-\mu,\mu]$, and denote by $\eta = \|n\|/\|F[m_*w_*]$
the noise-to-signal ratio. We show that any local minimizer $(m,w)$
of $\Ja$ satisfies 
\begin{itemize}
\item If $\eta \le 0.6$,
  \begin{equation}
    \label{eqn:mbnd}
    |m-m_*| \le (1+f(\eta))\frac{\lambda}{r},
  \end{equation}
  where $f(\eta)=2\eta+O(\eta^2)$ (Theorem \ref{thm:mnoiseres});
\item 
  $(m,{\bf 1}_{[-\lambda,\lambda]}w)$ is a solution of the FWI problem
  as stated above, that is,
  \begin{equation}
    \label{eqn:success}
    \lerr[m, {\bf 1}_{[-\lambda,\lambda]}w,d] \le \frac{1}{2}\epsilon^2,
  \end{equation}
  so long as
  \begin{eqnarray}
    \lambda & \ge & (2+f(\eta))\mu,\nonumber\\
    \epsilon &\ge & (4\pi r \alpha \lambda)^2 +\eta.
    \label{eqn:solcond}
  \end{eqnarray}
  (Theorem \ref{thm:ipnoisesuf})
\end{itemize}.
That is, the minimization of the extended objective $\Ja$ provides a
solution of the FWI problem, with constraints on the achievable
performance and bounds on the errors. All of these
bounds are shown to be sharp, and the constraints cannot be
significantly weakened. For example, the support radius $\lambda$ in
the projection onto $L^2([-\lambda,\lambda])$ must in general be at least twice the
support radius of the ``noise free'' wavelet $w_*$, plus an additional
allowance for noise.

The remainder of the paper begins with a complete description of the acoustic transmission problem and precise
statements of our main results, filling in the
few details not discussed in the preceding sketch. Then we provide an
discussion of the potential theoretical and practical import of these
results, including relations with prior work, some numerical
illustrations, and directions for further research. Following this
discussion section, we give complete proofs of all of our results.
An appendix describes the computational algorithm used in producing
the numerical examples, including a method for controlling the penalty
weight $\alpha$ based on the Discrepancy Principle
\cite[]{EnglHankeNeubauer,FuSymes2017discrepancy}. A second appendix
explains the relation between the support constraint and well-known
criteria for success of FWI based on data frequency content. The basis
for this relation is the Heisenberg inequality \cite[]{Folland:07}.

@@

We shall use the abbreviations ${\cal B}(X,Y)$ and ${\cal I}(X,Y)$ for 
the algebra of bounded linear operators from the Hilbert space $X$ to 
the Hilbert space $Y$, and its subalgebra of invertible operators.

\subsection{Acoustic transmission}
Assume small amplitude (linearized) acoustic propagation, constant
density, and isotropic point source and receiver. Denote by $m(\bx)$
the slowness (reciprocal velocity) at spatial position $\bx$, $w(t)$
the time dependence of the point source (``wavelet'') at location
$\bx=\bx_s$. Then the (excess) pressure field $p(\bx,t)$ obeys a
scalar wave equation:
\begin{eqnarray}
  \label{eqn:awe}
  \left(m(\bx)^2\frac{\partial^2 p}{\partial t^2} - \nabla^2\right) p(\bx,t) &=&
                                                                         w(t)\delta(\bx-\bx_s) \nonumber\\
  p(\bx,t)&=&0, t\ll 0
\end{eqnarray}
Suppose that a single trace is recorded, at distance $r>0$ from the
source position $\bx_s$. The dominant information in a single
trace is the transient signal time of arrival, constraining only the mean slowness in the
region
between source and receiver, so assume that the
slowness is constant, that is, independent of position $\bx$. The pressure field is simply the  the source
wavelet $w(t)$ convolved with the
acoustic Green's function, for which an analytic expression is
available in the constant $m$ case \cite[]{CourHil:62}:
\begin{equation}
  \label{eqn:homsol}
  p(\bx,t) = \frac{1}{4\pi |\bx-\bx_s|}w\left(t-m|\bx-\bx_s|\right).
\end{equation}

The receiver location $\bx_r$ lies at distance $r$ from the source
location $\bx_s$, that is, $|\bx_r-\bx_s|=r$. The predicted signal at
$p(\bx_r,t)$ depends nonlinearly on the slowness $m$ and linearly on the
source wavelet $w$. Therefore it is naturally represented as the
action of a $m$-dependent linear operator $F[m]$ on $w$, as described
in \ref{eqn:mod}.
%\begin{equation}
%\label{eqn:mod}
%F[m]w(t)= p(\bx_r,t) = \frac{1}{4\pi r}w\left(t-mr\right).
%\end{equation}

The slowness $m$ must be positive, as follows from basic acoustics,
and in fact reside in a range $M=(m_{\rm min}, m_{\rm max})$ characteristic of the
material model: for crustal rock, a reasonable choice would be
$m_{\rm min}=0.125, m_{\rm max}=0.6$ s/km.

Field recording takes place over finite time intervals (and with
finite sample rate, a detail we shall ignore in this paper). Denote by
$[t_{\rm min},t_{\rm max}]$ the recording interval for the
hypothetical single trace data. Then natural choices for domain and
range of $F[m]$ are $W = L^2(\bR)$ and $D=L^2([t_{\rm min},t_{\rm
  max}])$ respectively (that is, interpret the definition \ref{eqn:mod}
as holding for $t \in [t_{\rm min},t_{\rm  max}]$). So
\begin{equation}
  \label{eqn:mapprop}
  \mbox{for }m \in M, F[m] \in {\cal B}(W,D),\mbox{ and }\|F[m]\| =
  \frac{1}{4\pi r}.
\end{equation}
Note also that $F[m]$ is surjective for every $m \in M$.

In computational practice, $W$ will have to be replaced by a
finite-dimensional subspace of $L^2(\bR)$. Many such choices will
implicitly limit the support of $w \in W$ to a bounded interval, say
$[T_{\rm min},T_{\rm max}]$. To maintain the surjective property,
these bounds should be chosen so that $[t_{\rm min}, t_{\rm max}]
\subset [T_{\rm min}+mr,T_{\rm max}+mr]$ for all $m \in M$, that is,
\begin{eqnarray}
  \label{eqn:dombounds}
  t_{\rm min} &\ge & T_{\rm min}+m_{\rm max}r,\nonumber\\
  t_{\rm max} &\le & T_{\rm max}+m_{\rm min}r.
\end{eqnarray}
We will ignore these computational necessities in this work,
maintaining the definition $W=L^2(\bR)$.

\subsection{An inverse problem}

Part of the definition of an inverse problem based on the model
definition \ref{eqn:mod} is that observed data $d \in D$ be fit: that
is, $m, w$ must be found so that $F[m]w \approx d$. However this
formulation does not provide enough information to determine the model
$(m,w) \in M \times W$: indeed, the definition
\ref{eqn:mod} shows that $F[m]$ is surjective for any $m \in M$, so
the data can always be fit precisely by appropriate choice of $w$, and
$m$ is completely
unconstrained.  In order that data fit constrain $m$, it is necessary
to further constrain $w$ {\em a priori} - that is, to add information
beyond the trace data $d$.

One natural constraint is to assume that the support of $w$ is
limited. An assumption of small support may be justified as
follows. In practical seismic data processing, actual sources may act
over considerable time intervals - for example, the most commonly used
marine source (airgun) produces an oscillating pulse that dies away
slowly. It is commonplace to estimate this pulse then deconvolve it
from the data by safeguarded Fourier division or other means. This
so-called signature deconvolution results in modified data
corresponding to a source equal to signature deconvolution of the
pulse estimate itself. The deconvolved source pulse approximates an
impulse ($\delta(t)$): it cannot have point support, due to the finite
frequency nature of seismic data, hence the bandlimited nature of the
pulse estimate, but its support is ultimately much smaller than that
of the original pulse estimate.

Therefore, supplement the observed data $d \in D$ with a presumed 
source support radius $\lambda$. Define 
$\lW=\{w \in W: \mbox{ supp }w \subset [-\lambda,\lambda]\}$, and 
denote by $\lF$ the restriction of $F$, defined in equation 
\ref{eqn:mod}, to $M \times \lW$.

Note that for $w \in \lW$, $m \in M$,
$\mbox{supp }F[m]w \subset [ m_{\rm min}r-\lambda,
m_{\rm max}r+\lambda] \cap
[t_{\rm min},t_{\rm max}]$. Also, $\lF[m]$ is coercive for all $m \in
M$ only if
$[ m_{\rm min}r-\lambda,m_{\rm max}r+\lambda] \subset
[t_{\rm min},t_{\rm max}]$.

We will assume that $\lambda_{\rm max}
>0$ is chosen so that $M, r,\lambda_{\rm max}$ and $[t_{\rm min},t_{\rm max}]$ satisfy
\begin{equation}
  \label{eqn:fullrec}
  [ m_{\rm min}r-\lambda_{\rm max}, m_{\rm max}r+\lambda_{\rm max}]
  \subset [t_{\rm min},t_{\rm max}].
\end{equation}
Then for 
$m \in M$, $0 <\lambda \le \lambda_{\rm max}$, $\lF[m] \in {\cal B}(\lW,D)$ is coercive.

Note that condition \ref{eqn:fullrec} implies
\begin{eqnarray}
  \lambda_{\rm max} & \le & t_{\rm max}-m_{\rm max}r,\nonumber \\
  t_{\rm min}-m_{\rm min}r & \le & -\lambda_{\rm max} \label{eqn:lminc}
\end{eqnarray}
hence for any $m \in M$,
\begin{equation}
  [-\lambda_{\rm max}, -\lambda_{\rm max}] \subset[ t_{\rm min}-mr ,
  t_{\rm max}-mr].
  \label{eqn:zeroinc}
\end{equation}

@@

\noindent {\bf Problem Statement:} given data $d \in D$, relative 
error level $\epsilon \in [0,1)$, and support radius $\lambda \in (0, 
\lambda_{\rm max}]$, find 
$(m,w) \in M \times \lW$ for which 
\begin{equation}
  \label{eqn:probstat0}
  \|\lF[m]w-d\| \le \epsilon\|d\|. 
\end{equation}

@@

The best case for data fitting
is clearly the one in which the data can be fit precisely: that is,
there exists $(m_*,w_*) \in M\times \lW$ so that
\begin{equation}
  \label{eqn:defdatanonoise}
  d=\lF[m_*]w_*.
\end{equation}
Such data $d$ is {\em noise-free}, in the range of the map $\lF$. For
such data a solution of the Problem Statement
exists with arbitrarily small $\epsilon>0$.

\subsection{Soft constraints and penalty formulation}

As mentioned in the introduction, a straightforward nonlinear least
squares formulation of the problem stated in \ref{eqn:probstat0}
exhibits local minimizers with large fit error, and these will be
found by any local optimization algorithm unless it is initialized at
a quite accurate initial model. The first result stated in the next
section establishes this assertion for the single trace transmission
problem, however it appears to be valid for field scale problems, for
essentially the same reasons, as widely reported in the literature on
seismic inversion.

One (of many) alternative optimization formulation of this and similar
problems involves introducing an extension of the model-to-data
operator, and compensating for the physically spurious degrees of
freedom so introduced by adding a penalty term controlling them to the
objective.  The application of this idea presented here extends $\lF$
by dropping the support constraint on its second argument, that is,
using the expression \ref{eqn:mod} to define a map $F: M \times W
\rightarrow D$. The objective begins with an extended version of the
mean-square error,
\begin{equation}
  \label{eqn:mserr}
  e[m,w;d]=\frac{1}{2}\|F[m]w-d\|^2.
\end{equation}
We will choose a penalty to control the extra degrees of freedom, that
is, $w$ outside of $[-\lambda,\lambda]$. This  penalty is quadratic, constrains only $w\in W$ (as it is a soft
version of the condition $w \in \lW$), and is
$m-$independent. Therefore the penalty function takes the form
\begin{equation}
\label{eqn:pen}
p[w] = \frac{1}{2}\|Aw\|_Y^2 
\end{equation}
where $A \in {\cal B}(W,Y)$, $Y$ being another suitable Hilbert
space.

The penalty approach asks that a linear combination of $e,p$,
depending on a relative weight $\alpha$:
\begin{equation}
\Ja[m, w;d] = e[m,w;d] + \alpha^2 p[w] = \frac{1}{2}(\|F[m]w-d\|^2 + \alpha^2 \|Aw\|^2)
\label{eqn:pen}
\end{equation}
be minimized over $(m,w)\in M \times W$.

The choice of penalty weight $\alpha$ has a profound influence on the
character of this optimization problem. For the moment, we will mandate
only that $\alpha \ge 0$. A complete description of an algorithm using
the penalty function \ref{eqn:pen} to solve the inverse problem
\ref{eqn:probstat0} must include a selection principle for $\alpha$. We
describe one such principle in Appendix A.

>>>>

This paper examines one of a number of proposed alternatives to least
squares FWI, based on an extension of the model-to-data mapping.  We
show exactly how an example of this {\em extended inversion} approach
leads to solution of a very simple acoustic transmission
inverse problem {\em via} local gradient-based optimization. Despite
its simplicity, this problem exhibits several important
characteristics of all inverse problems in wave propagation in which
wave velocities are amongst the unknowns. It has the advantage of
permitting explicit calculations that reveal clearly the failure
mechanism for FWI, that is, straightforward nonlinear least squares,
and also the reasons for success of an extended inversion approach.

In the remainder of this introduction, we will briefly describe the conceptual
framework of this inverse problem, and sketch our main results. We leave out
some technical details that are supplied in the next section, where we give
precise statements of our results. In particular, theorems referenced
in the rest of this section are to be found, stated in full, in the
next section.

The physical setting for our discussion is the propagation of waves through a
homogeneous acoustic fluid occupying $\bR^3$ \cite[]{Frie:58}. The
fluid's salient characteristic for present purposes is the
slowness, or reciprocal wave velocity, $m > 0$. A source of acoustic
energy localized at a point in $\bR^3$ radiates uniformly in all
directions with time-dependent intensity $w(t)$. Both the acoustic
pressure field and the intensity (``wavelet'') $w$ vanish at large
negative time. A receiver records the pressure values at a distance
$r>0$ from the source position, between the times $t_{\rm min}$ and
$t_{\rm max}$. As explained in the next section, the
recorded presssure is predicted to be
\begin{equation}
\label{eqn:mod}
\frac{1}{4\pi r}w\left(t-mr\right)  = F[m]w(t)
\end{equation}
Ignoring the amplitude factor $1/(4\pi r)$, this map implements a $m$-dependent time
shift. This time shift operator is the basis of many illustrations of
the cycle-skipping phenomenon (for example, \cite{VirieuxOperto:09},
Figure 7), so it is unsurprising that an analysis of cycle-skipping
can be based on the simple modeling operator described above.

If the wavelet $w$ is square-integrable, so is the predicted
recorded pressure $F[m]w$ - its $L^2$ norm is the cumulative power trasferred
from the fluid to the receiver \cite[]{SantosaSymes:00}. Thus this relation defines a map
$F: \bR^+ \times L^2(\bR) \rightarrow
L^2([t_{\rm min},t_{\rm max}])$, linear in its second argument.

A naive version of the inverse problem investigated in this paper
would be: given recorded data
$d \in L^2([t_{\rm min},t_{\rm max}])$, find $m \in \bR^+$ and $w \in L^2(\bR)$ so that
$F[m]w \approx d$. Clearly this problem statement is not interesting as it
stands, since $F[m]$ is surjective for every $m \in \bR^+$: fitting
the data does not constrain the slowness $m$ at all.

To be useful, the problem statement must be augmented with a
constraint on the model $(m,w)$.  One natural constraint is to assume
that the support of $w$ is limited: choose $\lambda > 0$ and search
for $w$ with $\mbox{supp }w \subset [-\lambda,\lambda]$. The quantity
$\lambda$ is to be considered ``small''. Under some circumstances, it
is related to a mean wavelength of $w$ (Appendix B).

An assumption of small support may be justified as
follows. In practical seismic data processing, actual sources may act
over considerable time intervals. An example from exploration
seismology is the airgun, the most commonly used
marine source: an airgun (or array of airguns, the usual apparatus)
produces an oscillating pressure pulse that dies away
slowly. It is commonplace to estimate this pulse then deconvolve it
from the data by safeguarded Fourier division or other means. This
so-called signature deconvolution results in modified data
corresponding to a source equal to signature deconvolution of the
pulse estimate itself. The deconvolved source pulse approximates an
impulse ($\delta(t)$): it cannot have point support, due to the finite
frequency nature of seismic data, hence the bandlimited nature of the
pulse estimate, but its support is ultimately much smaller than that
of the original pulse estimate.

@@

Despite its simplicity, this problem exhibits the
cycle-skipping pathology characteristic of field-scale FWI. Specifically, we show that the reduced mean square
error function, the objective function of FWI for this problem, 
\begin{equation}
  \label{eqn:redms}
  \lerr[m,w;d]=\frac{1}{2}\|\lF[m]w-d\|^2/\|d\|^2,
\end{equation}
has local minimizers far from any global minimizer, even for noise-free
data (Theorem \ref{thm:fwi}).

This fact implies that local optimization must fail to solve a precise
version of the FWI problem: given $\epsilon, \lambda > 0$, $(m,w)$ is
a solution if it lies in the domain of $\lF$ and
$\|F[m]w-d\| \le \epsilon\|d\|$ (Problem Statement
\ref{eqn:probstat0} below).  In fact, we show
that``large residual'' local minima of $\lerr[\cdot,\cdot,d]$ exist,
at which its value is much greater than its global minimum
value. Therefore if $1/2\epsilon^2$ is near the global minimum value
of $\lerr[\cdot,cdot;d]$, then
local minimization of $\lerr[\cdot,\cdot;d]$ can fail to find a
solution of the inverse problem. This phenomenon is what is meant by
``cycle-skipping''.

Extended inversion involves an extension of the model-to-data map, in
this case $\lF$.  The extension used here consists in dropping the
support constraint on $w$, so the extended map is simply $F$ as
defined above ($\lF$ is a restriction of $F$, so $F$ is an extension
of $\lF$). To compensate for the resulting indeterminacy in the
estimation of $m$, we add a penalty term to the objective: for $m \in
\bR^+,w\in L^2(\bR), d \in L^2([t_{\rm min},t_{\rm max}])$, define
\begin{eqnarray}
  r[m,w;d] & = & \|F[m]w-d\|/\|d\|, \label{eqn:rmserr}\\
  e[m,w;d] & = & r[m,w;d]^2, \label{eqn:mserr}\\
  g[w] & = & \frac{1}{2}\|Aw\|^2/\|d\|^2, \label{eqn:pen} \\
  \Ja[m, w;d] &=& e[m,w;d] + \alpha^2 g[w] \nonumber\\
  &=& \frac{1}{2}(\|F[m]w-d\|^2 + \alpha^2 \|Aw\|^2)/\|d\|^2.
\label{eqn:penerr}
\end{eqnarray}

The normalizations by the data norm $\|d\|$ in the definitions
\ref{eqn:rmserr}, \ref{eqn:mserr}, \ref{eqn:pen}, and \ref{eqn:penerr} make the values
of $r$, $e$, $\alpha^2 g$, and $\Ja$ nondimensional.

The only obvious requirement on the penalty operator $A$ is that it
penalize 
nonzero values of $w$ for large $|t|$, so that minimizers $(m,w)$ of
$\Ja$ should have $w$
concentrated near $t=0$. Thus the penalty is a soft stand-in for the
support constraint. We
show that a good choice for $A$ is given (essentially) by
$(Aw)(t)=tw(t)$ (Proposition \ref{thm:rampgood}), and the
remainder of the discussion in this section makes this assumption. The
penalty weight $\alpha$ must be chosen somehow; we review a simple
algorithm for controlling $\alpha$ later in the paper.

Assume that $\mu>0$ and $d = F[m_*]w_* + n$, with $\mbox{supp }w_*
\subset [-\mu,\mu]$, and denote by $\eta = \|n\|/\|F[m_*,w_*]\|$
the noise-to-signal ratio. We show that any local minimizer $(m,w)$
of $\Ja$ satisfies 
\begin{itemize}
\item If $\eta \le 0.6$,
  \begin{equation}
    \label{eqn:mbnd}
    |m-m_*| \le (1+f(\eta))\frac{\mu}{r},
  \end{equation}
  where $f(\eta)=2\eta+O(\eta^2)$ (Theorem \ref{thm:mnoiseres});
\item 
  $(m,{\bf 1}_{[-\lambda,\lambda]}w)$ is a solution of the FWI problem
  as stated above, that is,
  \begin{equation}
    \label{eqn:success}
    \lerr[m, {\bf 1}_{[-\lambda,\lambda]}w;d] \le \frac{1}{2}\epsilon^2,
  \end{equation}
  so long as
  \begin{eqnarray}
    \lambda & \ge & (2+f(\eta))\mu,\nonumber\\
    \epsilon &\ge & (4\pi r \alpha \lambda)^2 +\eta.
    \label{eqn:solcond}
  \end{eqnarray}
  (Theorem \ref{thm:ipnoisesuf})
\end{itemize}.
That is, the minimization of the extended objective $\Ja$ provides a
solution of the FWI problem, with constraints on the achievable
performance and bounds on the errors. All of these
bounds are shown to be sharp, and the constraints cannot be
significantly weakened. For example, in conclusion \ref{eqn:solcond}, the support radius $\lambda$ in
the projection onto $L^2([-\lambda,\lambda])$ must in general be at least twice the
support radius of the ``noise free'' wavelet $w_*$, plus an additional
allowance for noise.

The remainder of the paper begins with a complete description of the acoustic transmission problem and precise
statements of our main results, filling in the
few details not discussed in the preceding sketch. Then we provide a
discussion of the potential theoretical and practical import of these
results, including relation with prior work, some numerical
illustrations, and directions for further research. Following this
discussion section, we give complete proofs of all of our results.
An appendix describes the computational algorithm used in producing
the numerical examples, including a method for controlling the penalty
weight $\alpha$ based on the Discrepancy Principle
\cite[]{EnglHankeNeubauer,FuSymes2017discrepancy}. A second appendix
explains the relation between the support constraint and well-known
criteria for success of FWI based on data frequency content. The basis
for this relation is the Heisenberg inequality \cite[]{Folland:07}.

@@

We shall use the abbreviations ${\cal B}(X,Y)$ and ${\cal I}(X,Y)$ for 
the algebra of bounded linear operators from the Hilbert space $X$ to 
the Hilbert space $Y$, and its subalgebra of invertible
operators. All of the operators appearing in this discussion are
members of ${\cal B}(X,Y)$ for some choice of Hilbert spaces $X$ and
$Y$, and superscript $T$ denotes the transpose or adjoint of a such an
operator in the sense specified by its domain and range Hilbert structure.

\subsection{Acoustic transmission}
Assume small amplitude (linearized) acoustic propagation, constant
density, and isotropic point source and receiver. Denote by $m(\bx)$
the slowness (reciprocal velocity) at spatial position $\bx$, $w(t)$
the time dependence of the point source intensity (``wavelet'' or
``source wavelet'') at location
$\bx=\bx_s$. Then the (excess) pressure field $p(\bx,t)$ obeys a
scalar wave equation \cite[]{Frie:58}:
\begin{eqnarray}
  \label{eqn:awe}
  \left(m(\bx)^2\frac{\partial^2 p}{\partial t^2} - \nabla^2\right) p(\bx,t) &=&
                                                                         w(t)\delta(\bx-\bx_s) \nonumber\\
  p(\bx,t)&=&0, t\ll 0
\end{eqnarray}
Suppose that the pressure at a receiver position $\bx=\bx_r$ is recorded, at distance $r>0$ from the
source position $\bx_s$. We shall refer to the resulting function of
time as a {\em trace}, from common seismic usage. The dominant information in a single
trace is the transient signal time of arrival, constraining only the mean slowness in the
region
between source and receiver, so assume that the
slowness is constant, that is, independent of position $\bx$. The pressure field is simply the  the source
wavelet $w(t)$ convolved with the
acoustic Green's function, for which an analytic expression is
available in the constant $m$ case \cite[]{CourHil:62}:
\begin{equation}
  \label{eqn:homsol}
  p(\bx,t) = \frac{1}{4\pi |\bx-\bx_s|}w\left(t-m|\bx-\bx_s|\right).
\end{equation}

Define $r=|\bx_r-\bx_s| $ to be the source-receiver distance
(``offset''). The predicted signal at
$p(\bx_r,t)$ depends nonlinearly on the slowness $m$ and linearly on the
source wavelet $w$. Therefore it is naturally represented as the
action of a $m$-dependent linear operator $F[m]$ on $w$, as described
in \ref{eqn:mod}, which we repeat here for convenience:
\begin{equation}
\label{eqn:mod1}
F[m]w(t)= p(\bx_r,t) = \frac{1}{4\pi r}w\left(t-mr\right).
\end{equation}

The slowness $m$ must be positive, as follows from basic acoustics,
and in fact reside in a range characteristic of the
material model: for crustal rock, a reasonable choice would be
$m_{\rm min}=0.125, m_{\rm max}=0.6$ s/km.

Field recording takes place over finite time intervals (and with
finite sample rate, a detail we shall ignore in this paper). Denote by
$[t_{\rm min},t_{\rm max}]$ the recording interval for the
hypothetical single trace data.

Natural choices for domain and
range of $F$ are
\begin{itemize}
\item $M=(m_{\rm min}, m_{\rm max}),\,0 < m_{\rm min} \le m_{\rm
    max}$;
\item $W = L^2(\bR)$;
\item $D=L^2([t_{\rm min},t_{\rm max}]),\, t_{\rm min}<t_{\rm max}$;
\item $F: M \times W \rightarrow D$ as specified in \ref{eqn:mod}.
\end{itemize}
It is immediately evident from these choices and from the definition
\ref{eqn:mod} that
\begin{equation}
  \label{eqn:mapprop}
  \mbox{for }m \in M, F[m] \in {\cal B}(W,D),\mbox{ and }\|F[m]\| =
  \frac{1}{4\pi r}.
\end{equation}
Note also that $F[m]$ is surjective for every $m \in M$.

\noindent{\bf Remark:} In computational practice, $W$ will have to be replaced by a
finite-dimensional subspace of $L^2(\bR)$. Many such choices will
implicitly limit the support of $w \in W$ to a bounded interval, say
$[T_{\rm min},T_{\rm max}]$. To maintain the surjective property,
these bounds should be chosen so that $[t_{\rm min}, t_{\rm max}]
\subset [T_{\rm min}+mr,T_{\rm max}+mr]$ for all $m \in M$, that is,
\begin{eqnarray}
  \label{eqn:dombounds}
  t_{\rm min} &\ge & T_{\rm min}+m_{\rm max}r,\nonumber\\
  t_{\rm max} &\le & T_{\rm max}+m_{\rm min}r.
\end{eqnarray}
We will ignore these computational necessities in this work,
maintaining the definition $W=L^2(\bR)$.

\subsection{An inverse problem}
As mentioned earlier, $F[m]$ is surjective for every
$m \in M$. Since all possible data lie in the range of $F[m]$ for any
$m \in M$, some restriction of the domain
of $F$ is necessary in order that fitting the data constrain
$m$. A more precise version of the constraint proposed in the
introduction involves a selection of a maxium support radius
$\lambda_{\rm max} >0$. Then define for $\lambda \in (0,\lambda_{\rm max}]$:
\begin{itemize}
\item $\lW = \{w \in W:
  \mbox{supp }w \subset [-\lambda,\lambda]\}$;
\item $\lF = F|_{M \times \lW}$.
\end{itemize}

@@

\begin{quote}
\noindent {\bf Problem Statement:}
  given data $d \in D$, relative error level $\epsilon \in
  [0,1)$, and support radius $\lambda \in (0, \lambda_{\rm
    max}]$, find $(m,w) \in M \times \lW$ for which 
\begin{equation}
  \label{eqn:probstat0}  \|\lF[m]w-d\| \le \epsilon\|d\|,
\end{equation}
or equivalently
\begin{equation}
  \label{eqn:probstat1}
  \lerr[m,w;d] \le \frac{1}{2}\epsilon^2,
\end{equation}
with $\lerr$ defined in equation \ref{eqn:redms}.
\end{quote}

@@

The best case for data fitting
is clearly the one in which the data can be fit precisely: that is,
there exists $(m_*,w_*) \in M\times \lW$ so that
\begin{equation}
  \label{eqn:defdatanonoise}
  d=\lF[m_*]w_*.
\end{equation}
Such data $d$ is {\em noise-free}, in the range of the map $\lF$. For
such data a solution of the Problem Statement \ref{eqn:probstat0}
exists with arbitrarily small $\epsilon>0$.

\subsection{Main Results}
While $F[m]$ is surjective for every $m \in M$, as noted above, it is
very far from injective. On the other hand, under a constraint that
will be assumed throughout, $\lF[m]$ is injective (in fact, $4 \pi r
\lF[m]$ is an isometry):
\begin{proposition}
  \label{thm:fullrec}
  Suppose that 
  \begin{equation}
    \label{eqn:fullrec}
    [ m_{\rm min}r-\lambda_{\rm max}, m_{\rm max}r+\lambda_{\rm max}]
    \subset [t_{\rm min},t_{\rm max}].
  \end{equation}
  Then $\lF[m]\in {\cal B}(\lW,D)$ is coercive for every $m \in M, \lambda \in
  (0,\lambda_{\rm max}]$.
\end{proposition}

The first main result establishes the existence of large (100 \%)
residual local minimizers for the basic FWI objective $\lerr$, even
for noise-free data.
\begin{theorem}
  \label{thm:fwi}
  Suppose that $0 <\lambda\le \lambda_{\rm max}$,  $m_* \in M, w_*
  \in \lW, d=\lF[m_*]w_*$ is noise-free data per definition \ref{eqn:defdatanonoise},
  Under assumption \ref{eqn:fullrec}, for any $m \in M$ with $|m-m_*|r>2\lambda$,
\begin{equation}
  \label{eqn:isovpm}
 \min_w \lerr[m,w;d]=\lerr[m,0;d] = \frac{1}{2},
\end{equation}
and any such $(m,0)$ is a local minimizer of $\lerr$ with relative RMS
error = 1.0.
\end{theorem}

The main theoretical device used in the proofs of our main results on 
extended inversion is the reduction of the penalty objective $\Ja$
(equation \ref{eqn:penerr}) to a function $\tJa$ of 
$m$ alone, by minimization over $w$: 
\begin{equation}
  \label{eqn:redexp}
  \tJa[m;d] = \inf_w \Ja[m,w;d]
\end{equation}
This reduction is the foundation 
of the {\em Variable Projection Method} (VPM) 
\cite[]{GolubPereyra:03} for solution of partly linear (``separable'') 
least squares problems.
